---
title: "Stats Assignment"
date: 22-07-2017
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(ggplot2)
library(gridExtra)
library(extrafont)

# Import data file from csv and clean data
data <- read.csv('question1CSV.csv', header=T)

```

## Question 1
### Summary Table for Study Participants Risk Factors
**Factor** | **Summary**
------------------------|:-----------------:
**Age**, years (range, mean (sd))   | `r min(data$Age)`-`r max(data$Age)`, `r round(mean(data$Age), 2)` (`r round(sd(data$Age), 2)`)
**Gender**, n (%)                         |
  *Male*                                  | `r summary(data$Gender)['male']` (`r round(100*summary(data$Gender)['male']/length(data$Gender), 1)`)
  *Female*                                | `r summary(data$Gender)['female']` (`r round(100*summary(data$Gender)['female']/length(data$Gender), 1)`)
**Diabetes**, n (%)                       | `r summary(data$Diabetes)['yes']` (`r round(100*summary(data$Diabetes)['yes']/length(data$Diabetes), 1)`)
**Smoker**, n (%)                         | `r summary(data$Ever.smoked)['yes']` (`r round(100*summary(data$Ever.smoked)['yes']/length(data$Ever.smoked), 1)`)
**Peripheral Vascular Disease**, n (%)    | `r summary(data$PVD)['yes']` (`r round(100*summary(data$PVD)['yes']/length(data$PVD), 1)`)
**Cerebrovascular Disease**, n (%)        | `r summary(data$CVD)['yes']` (`r round(100*summary(data$CVD)['yes']/length(data$CVD), 1)`)
**Hypercholestrolemia**, n (%)            | `r summary(data$Hypertension)['yes']` (`r round(100*summary(data$Hypertension)['yes']/length(data$Hypertension), 1)`)

### Comparison of Indicies for Severity of Disease in Radial Artery for Smokers and Non-Smokers
```{r part(b), echo=FALSE} 
# Generate three boxplots to compar main indicies for disease severity for radial artery depending on whether patient has ever smoked.
# plot1 - radial artery luminal narrowing depending on smoker
# plot2 - radial artery intimal thickness index depending on smoker
# plot3 - radial artery intima to media ratio depending on smoker

    plot1 <- ggplot(data, aes(Ever.smoked,RA.Luminal.narrowing....)) + 
    geom_boxplot() + 
    coord_flip() +
    theme_bw(base_size = 8) +
    scale_x_discrete(labels=c("No","Yes")) +
    scale_y_continuous(breaks = seq(0, 70, 10), limits=c(0,70), expand=c(0,0)) +
    labs(
      y="Luminal Narrowing (%)", 
      x="Smoker?")
  ggsave("plot1.1.tiff", height=4.6, width=17, units="cm", dpi=600)
  
  plot2 <- ggplot(data, aes(Ever.smoked,RA.Intimal.thickness.index)) + 
    geom_boxplot() + 
    coord_flip() +
    theme_bw(base_size = 8) +
    scale_x_discrete(labels=c("No","Yes")) +
    scale_y_continuous(breaks = seq(0, 2, 0.2), limits=c(0,2), expand=c(0,0)) +
    labs(
      y="Intimal Thickness Index", 
      x="Smoker?")
  ggsave("plot1.2.tiff", height=4.6, width=17, units="cm", dpi=600)
  
  plot3 <- ggplot(data, aes(Ever.smoked,RA.Intima.to.media.ratio)) + 
    geom_boxplot() + 
    coord_flip() +
    theme_bw(base_size = 8) +
    scale_x_discrete(labels=c("No","Yes")) +
    scale_y_continuous(breaks = seq(0, 11, 1), limits=c(0,11), expand=c(0,0)) +
    labs(
      y="Intima-to-Media Ratio", 
      x="Smoker?")
  ggsave("plot1.3.tiff", height=4.6, width=17, units="cm", dpi=600)
  
  
  blank <- ggplot() + theme_void()
  
  # Using gridExtra package to generate all three plots into one image
  grid.arrange(plot1, blank, plot2, blank, plot3, heights=c(0.30, 0.05, 0.30, 0.05, 0.3), ncol=1)
  
```


```{r ttests, echo=FALSE}
  # part (c) - boxplot display of 
  ttest1 <- t.test(data$RA.Luminal.narrowing.... ~ data$Ever.smoked, var.equal=T, conf.level=0.95)
  ttest2 <- t.test(data$RA.Intimal.thickness.index ~ data$Ever.smoked, var.equal=T, conf.level=0.95)
  ttest3 <- t.test(data$RA.Intima.to.media.ratio ~ data$Ever.smoked, var.equal=T, conf.level=0.95)
```  

### Difference between Smokers and Non-Smokers on each Index of Disease Severity for Radial Artery
**Two Sample t-test: Non-smoker - Smoker**

Disease Severity Indicies  | Difference | 95% CI  | t-value | p-value
-------------------------- | ------------------ | ----------- | -------- | -------
Luminal Narrowing (%)      | `r round(ttest1$estimate[1]-ttest1$estimate[2],3)` | `r round(ttest1$conf.int,3)` | `r round(ttest1$statistic,3)` | `r round(ttest1$p.value,3)`
Intimal Thickness Index    | `r round(ttest2$estimate[1]-ttest2$estimate[2],3)` | `r round(ttest2$conf.int,3)` | `r round(ttest2$statistic,3)` | `r round(ttest2$p.value,3)`  
Intima-to-Media Ratio      | `r round(ttest3$estimate[1]-ttest3$estimate[2],3)` | `r round(ttest3$conf.int,3)` | `r round(ttest3$statistic,3)` | `r round(ttest3$p.value,3)` 


-------------------------------------------------------------
                         Mean (st. dev)
                         Smoker
----------------------- ---------------- --------------- -------------------------
Luminal Narrowing (%)   
Intima Thickness Index
Intima-to-Media Ratio
-------------------------------------------------------------
```{r table}

differences <- data.frame(x = c('Luminal Narrowing (%)', 'Intima Thickness Index', 'Intima-to-Media Ratio'), min=c(-0.164, -0.053, -0.607), max=c(8.766, 0.119, 0.502), diff=c(4.301, 0.033, -0.053))

plot4 <- ggplot(differences, aes(x, diff)) + 
         geom_hline(yintercept=0, size=0.3) +
         geom_point(aes(x, diff, color="red")) +
         geom_errorbar(aes(x, ymin = min, ymax = max, color="red")) +
         coord_flip() +
         theme_bw(base_size = 8) +
         scale_y_continuous(breaks = seq(-1, 9, 1), limits=c(-1,9), expand=c(0,0)) +
         guides(color=FALSE) +
         labs(
           y='Difference in Means (Smoker - Non-Smoker)', 
           x='Indicies for Disease Severity')

plot4
ggsave("plot1.4.tiff", height=5.9, width=15.73, units="cm", dpi=600)

```